In this report, the design and implementation of a automated digital chip tester is presented. The work was carried out as the group design project of \textit{Team I} for module \emph{ELEC6027 - VLSI Design Project} in the second semester of the \textit{MSc System-on-Chip / European Masters in Embedded Computing Systems} in the University of Southampton.

The aim of the project was to design and implement an automated tester for testing of the University of Southampton Superchip, designed by undergraduate students in the University of Southampton as part of their course syllabus. The tester must carry out verification of the functionality of the chips and provide a feedback on the success of the design or the errors in this functionality.

For the VLSI project, the teams were allocated with \textit{Altera DE2-115} development boards on which to implement their designs. Additionally, the teams were allocated a budget to make amands for extra components they need, such as external PCBs or ICs.

The implementation presented in this report consists of an SoPC (System-on-Programmable-Chip) programmed on the DE2 board that interfaces with two external PCBs that serve as hosts to the DUT (Device-Under-Test) - namely the Southampton Superchip - and a slave FPGA that provides virtual designs for testing. Additionally, the main DE2 board interfaces with a database server that holds a database to hold information and results from tests, as well as a backend to provide a user-friendly GUI.

The design presented in this paper has been successfully completed within the second semester of the academic year $2011/12$, providing the required functionality. The design was completed keeping performance and power consumption in mind, and also focusing on efficiency, correctness of testing and reliability, as well as security and user-friendliness.